### 2020-03-24 ChapterBackground/chapterBackground.tex
Updated: Removed all subsubsections

\subsubsection{Client-Server}
Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players \cite{Ferretti2018}. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the sate of the central server, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PForMMOs}.

@@

\subsubsection{Peer-to-Peer}
Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as both server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of short-session-based games are lower than persistent games, making peer-to-peer architectures better suited to short-session-based games \cite{P2PForMMOs}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PForMMOs}.

@@

\subsubsection{Hybrid Peer-to-Peer}
Hybrid Peer-to-Peer architectures are a combination of client-server and peer-to-peer architectures. The peer-to-peer part of the system can be responsible for different aspects of the system, including Cooperative Message Dissemination, State Distribution, Basic Server Control and distributing software updates. \cite{P2PForMMOs}. Hybrid Peer-to-Peer enables trade-offs between consistency control and scalability, determined by how the system is divided between server and peer-to-peer. The more responsibility peers have, the greater the scalability but at a loss of consistency control.
%These can be broadly classified into two types: Homogeneous, in which all nodes have similar responsibilities and Heterogeneous, in which some peers be

\subsubsection{Distributed server}
% TODO: More references for this
Distributed server or Multi-Server architectures use multiple servers and the computational workload is divided between the servers. This can be achieved in one of two ways, either through `shards' or through division of the game world.

@@

Interest Management can be broadly classified into two categories: Regions and Auras \cite{Morgan2005, storey2004determining}.
\subsubsection{Regions}
``In the region based approach the virtual world is commonly, but not always, divided into well defined uniform sized regions that are static in nature (i.e. their boundaries are defined at virtual world creation time)."\cite{storey2004determining}
\subsubsection{Auras}
``In the aura based approach each object is associated to an aura that defines an area of the virtual world over which an object may exert influence. Ideally, an object may potentially communicate their actions to only objects that fall within their influence."\cite{storey2004determining}\\

>>>>

Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players \cite{Ferretti2018}. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the sate of the central server, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PForMMOs}.

@@

Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as both server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of short-session-based games are lower than persistent games, making peer-to-peer architectures better suited to short-session-based games \cite{P2PForMMOs}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PForMMOs}.

@@

Hybrid Peer-to-Peer architectures are a combination of client-server and peer-to-peer architectures. The peer-to-peer part of the system can be responsible for different aspects of the system, including Cooperative Message Dissemination, State Distribution, Basic Server Control and distributing software updates. \cite{P2PForMMOs}. Hybrid Peer-to-Peer enables trade-offs between consistency control and scalability, determined by how the system is divided between server and peer-to-peer. The more responsibility peers have, the greater the scalability but at a loss of consistency control.
%These can be broadly classified into two types: Homogeneous, in which all nodes have similar responsibilities and Heterogeneous, in which some peers be

% TODO: More references for this
Distributed server or Multi-Server architectures use multiple servers and the computational workload is divided between the servers. This can be achieved in one of two ways, either through `shards' or through division of the game world.

@@

Interest Management can be broadly classified into two categories: Regions and Auras \cite{Morgan2005, storey2004determining}.

``In the region based approach the virtual world is commonly, but not always, divided into well defined uniform sized regions that are static in nature (i.e. their boundaries are defined at virtual world creation time)."\cite{storey2004determining}\\

``In the aura based approach each object is associated to an aura that defines an area of the virtual world over which an object may exert influence. Ideally, an object may potentially communicate their actions to only objects that fall within their influence."\cite{storey2004determining}\\

### 2020-03-24 ChapterResults/chapterResults.tex
Updated: Removed all subsubsections

\subsubsection{Experiment parameters and output}
The two servers use the following tolerances unless otherwise stated: a maximum speed tolerance of $32m\mathord{\cdot}s^{-1}$; maximum latency tolerance of $2ms$; and a maximum frame-time tolerance of $15ms$. Each server has a random delay before starting between $0$ and the target frame-time in order to prevent the two servers always having synchronous update cycles. Each experiment is repeated $50$ times. An experiment was conducted for each aura tolerance value. The aura size for the varying tolerances are plotted and the tolerances used are marked. For each experiment, three tolerances values were chosen. For each tolerance value, the following output graphs were plotted:
\begin{itemize}
	\item The ratio of late to correct collisions and ratio of misses to total collision.
	\item The mean and $+/-2$ standard deviations of penetration time were plotted against each tolerance factor for each of the three tolerance values chosen.
\end{itemize}

@@

\subsubsection{Varying Speed}

In this experiment the speed of each of the two objects was varied from $1m\mathord{\cdot}s^{-1}$ to double the the following tolerances: $16m\mathord{\cdot}s^{-1}$; $32m\mathord{\cdot}s^{-1}$ and $48m\mathord{\cdot}s^{-1}$. Increments of $1m\mathord{\cdot}s^{-1}$ were used and the whole experiment repeated $50$ times.

@@

\subsubsection{Varying Latency}

In this experiment the latency between the two servers was varied from $0ms$ to double the following tolerances: $2ms$; $8ms$ and $16ms$ in increments of $0.05ms$, $0.25ms$ and $1ms$ respectively. The latency used is the target latency, in reality the latency will never exactly be equal to $0ms$.

@@

\subsubsection{Varying Frame-Time}

In this experiment the frame-time was varied from $0ms$ to double the frame-time tolerance for the following frame-times: $8.33ms$ $(120Hz)$; $15ms$ $(66.67Hz)$ and $33.33ms$ $(30Hz)$. Increments of $1ms$, $1ms$ and $2ms$ were used respectively. The frame-time used is target frame time, in reality frame-time cannot be equal to exactly $0ms$.

@@

\subsubsection{Thrashing}
In addition to recording the number of collisions that resulted in thrashing, the number of times an object thrashed between servers was also measured. The number of times thrashed is determined by measuring the number of migrations away from the server that the collisions takes place on, before the collision takes place. A histogram of the number of occurrences against the number of times thrashed is shown in Fig. \ref{fig_ThrashingHistogram}. For this histogram, data from the Frame-Time $33.33ms$ experiment is used, as this provides the largest sample of thrashing. Fig. \ref{fig_ThrashingHistogram} demonstrates that the most frequent number of times thrashed is 1, by a large margin followed by a downward trend.

@@

\subsubsection{Evaluation - Speed}
As the speed increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsSpeedLow}, \ref{fig_RatioVsSpeedMid} and \ref{fig_RatioVsSpeedHigh}. When speed is below the speed tolerance, late collisions do not exceed 10\%. This 10\% is assumed to be caused by false-positives. Once above the speed tolerance, the proportion of late collisions increases immediately, which is predicted by the aura calculation. Missed collisions begin to occur when speed exceed $80ms^{-1}$ in the $48ms^{-1}$ experiment, this is due to objects travelling so quickly, the objects completely pass one another, missing the collision entirely.

@@

\subsubsection{Evaluation - Latency}
As the latency increases, the proportion of late collisions to correct collisions increases, as shown in Fig. \ref{fig_RatioVsLatencyHigh}. However, Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyLow} do not show this relationship. In the $2ms$ latency tolerance experiment, late collisions do not exceed $20\%$, which is assumed to be accounted for by false-positives. The aura calculation predicts that late collisions should start occurring once $3ms$ has been exceeded, however, this is not observed, as the proportion does not increase above what would be expected for false-positives.
In the $8ms$ latency tolerance experiment, late collisions do not exceed $2\%$, which is assumed to be accounted for by false-positives. This remains the case even when the tolerance is exceeded. The aura calculation predicts that late collisions will not begin to occur until $20ms$ latency is reached, which this experiment does not reach, therefore no late collisions outside of false-positives are expected. In the $16ms$ latency tolerance experiment, late collisions do not exceed $3\%$ below the tolerance, which is assumed to be accounted for by false-positives. The aura calculation predicts that late collisions will not begin to occur until $20ms$ latency is reached. In Fig. \ref{fig_RatioVsLatencyHigh} we see the proportion of errors increase from $20ms$ and above, however, it does not exceed the expected false-positive proportion until $25ms$ latency is reached.
%TODO: Re-run latency experiment, larger sample size?

@@

\subsubsection{Evaluation - Frame-Time}
As frame-time increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyMid}, however in \ref{fig_RatioVsLatencyHigh} the sample size is small above the tolerance line due to the high occurrences of thrashing and so the relationship in this experiment can not be deduced. For the $8ms$ and $15ms$ experiments the late collisions do not exceed 10\% and are assumed to be accounted for by false-positives. In the $8ms$ frame-time tolerance experiment, late collisions are not expected to occur until $16ms$ frame-time, based on the aura size. Late collisions do not exceed $20\%$ and is assumed to be accounted for by false-positives. In the $15ms$ frame-time tolerance experiment, late collisions exceed $20\%$ once $16ms$ latency is exceeded as predicted by the aura size and increase to over $75\%$ at $30ms$ latency.
In the final frame-time tolerance experiment using $32ms$ frame-time tolerance, late collisions only occur for frame-times of $64ms$. The aura calculation predicts that late collisions should begin to occur as soon as $32ms$ frame-time is exceeded, however, due to the low sample size, as a result of the high occurrences of thrashing, the plot does not accurately represent any relationship between frame-time and proportion of late collisions.

@@

\subsubsection{Varying Factors - Summary}
The varying factors experiments demonstrate that the aura calculation is conclusively correct for speed and frame-time aura tolerance factors and at minimum approximate for latency. For each varied factor, for all three tolerance values chosen, below the tolerance value, the mean + two standard deviations lies under the maximum penetration time expected on a centralised simulation, with only a small number of exceptions discussed above. These experiments also demonstrate the relationship between increasing the tolerance values and amount of thrashing expected to occur. There are no missed collisions, with the exception of the speed experiments and these have significantly fewer than $1\%$ misses below the tolerance and $1.12\%$ misses above the tolerance. Once factors increase above their respective tolerance values, late collisions begin to occur and the more the factors are exceeded, the number and magnitude of errors increase. This demonstrates the correctness of the aura calculation in AP.
%Therefore, we have shown that if the aura calculation factors remain within their respective tolerances, erroneous collisions will be small in magnitude and rare in number. %TODO: Summary Table: Row for each factor: Number of collisions below tolerance, Percentage of errors below tolerance, Mean error below tolerance

@@

\subsubsection{Experiment parameters and output}
The two servers use the following tolerances: a maximum speed tolerance of $32m\mathord{\cdot}s^{-1}$; maximum latency tolerance of $2ms$; and a maximum frame-time tolerance of $15ms$. Each server has a random delay before starting between $0$ and the target frame-time in order to prevent the two servers always having synchronous update cycles. Each experiment is repeated $50$ times. The mean and $+/-2$ standard deviations of penetration time were plotted against each tolerance factor for each of the three tolerance values chosen. In addition, the ratio of errors to correct collisions and misses to total collisions were plotted against packet-loss.

@@

\subsubsection{Evaluation}
As packet-loss increases, late and erroneous collisions increase and late collisions increase in magnitude. This is due to messages being delayed as packets are lost and need to be re-sent by RakNet's reliability layer. Missed collisions begin to occur as soon as packet-loss exceeds $1\%$, but increase at a lower rate relative to late collisions. (Note, late collisions begin to occur from $0\%$, this is expected as all factors are set to be equal to their tolerances). These results demonstrate that AP is able to prevent $50\%$ of late collisions even in cases of packet-loss as high as (15\%) and prevent all missed collisions with up to $1\%$ packet-loss, which is much greater packet-loss than would be expected in practice (0.01\%) \cite{ThousandEyesCloudPerf2018}.

>>>>

For the varying factors experiments, the following parameters were used and output given.
The two servers use the following tolerances unless otherwise stated: a maximum speed tolerance of $32m\mathord{\cdot}s^{-1}$; maximum latency tolerance of $2ms$; and a maximum frame-time tolerance of $15ms$. Each server has a random delay before starting between $0$ and the target frame-time in order to prevent the two servers always having synchronous update cycles. Each experiment is repeated $50$ times. An experiment was conducted for each aura tolerance value. The aura size for the varying tolerances are plotted and the tolerances used are marked. For each experiment, three tolerances values were chosen. For each tolerance value, the following output graphs were plotted:
\begin{itemize}
	\item The ratio of late to correct collisions and ratio of misses to total collision.
	\item The mean and $+/-2$ standard deviations of penetration time were plotted against each tolerance factor for each of the three tolerance values chosen.
\end{itemize}

@@

The first experiment conducted was the varying speed experiment. In this experiment the speed of each of the two objects was varied from $1m\mathord{\cdot}s^{-1}$ to double the the following tolerances: $16m\mathord{\cdot}s^{-1}$; $32m\mathord{\cdot}s^{-1}$ and $48m\mathord{\cdot}s^{-1}$. Increments of $1m\mathord{\cdot}s^{-1}$ were used and the whole experiment repeated $50$ times.

@@

The second experiment conducted was the varying latency experiment. In this experiment the latency between the two servers was varied from $0ms$ to double the following tolerances: $2ms$; $8ms$ and $16ms$ in increments of $0.05ms$, $0.25ms$ and $1ms$ respectively. The latency used is the target latency, in reality the latency will never exactly be equal to $0ms$.

@@

The third experiment conducted was the varying frame-time experiment. In this experiment the frame-time was varied from $0ms$ to double the frame-time tolerance for the following frame-times: $8.33ms$ $(120Hz)$; $15ms$ $(66.67Hz)$ and $33.33ms$ $(30Hz)$. Increments of $1ms$, $1ms$ and $2ms$ were used respectively. The frame-time used is target frame time, in reality frame-time cannot be equal to exactly $0ms$.

@@

In addition to recording the number of collisions that resulted in thrashing, the number of times an object thrashed between servers was also measured. The number of times thrashed is determined by measuring the number of migrations away from the server that the collisions takes place on, before the collision takes place. A histogram of the number of occurrences against the number of times thrashed is shown in Fig. \ref{fig_ThrashingHistogram}. For this histogram, data from the Frame-Time $33.33ms$ experiment is used, as this provides the largest sample of thrashing. Fig. \ref{fig_ThrashingHistogram} demonstrates that the most frequent number of times thrashed is 1, by a large margin followed by a downward trend.

@@

\subsection{Experiment- Varying Factors Evaluation}
As the speed increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsSpeedLow}, \ref{fig_RatioVsSpeedMid} and \ref{fig_RatioVsSpeedHigh}. When speed is below the speed tolerance, late collisions do not exceed 10\%. This 10\% is assumed to be caused by false-positives. Once above the speed tolerance, the proportion of late collisions increases immediately, which is predicted by the aura calculation. Missed collisions begin to occur when speed exceed $80ms^{-1}$ in the $48ms^{-1}$ experiment, this is due to objects travelling so quickly, the objects completely pass one another, missing the collision entirely.

@@

As the latency increases, the proportion of late collisions to correct collisions increases, as shown in Fig. \ref{fig_RatioVsLatencyHigh}. However, Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyLow} do not show this relationship. In the $2ms$ latency tolerance experiment, late collisions do not exceed $20\%$, which is assumed to be accounted for by false-positives. The aura calculation predicts that late collisions should start occurring once $3ms$ has been exceeded, however, this is not observed, as the proportion does not increase above what would be expected for false-positives.
In the $8ms$ latency tolerance experiment, late collisions do not exceed $2\%$, which is assumed to be accounted for by false-positives. This remains the case even when the tolerance is exceeded. The aura calculation predicts that late collisions will not begin to occur until $20ms$ latency is reached, which this experiment does not reach, therefore no late collisions outside of false-positives are expected. In the $16ms$ latency tolerance experiment, late collisions do not exceed $3\%$ below the tolerance, which is assumed to be accounted for by false-positives. The aura calculation predicts that late collisions will not begin to occur until $20ms$ latency is reached. In Fig. \ref{fig_RatioVsLatencyHigh} we see the proportion of errors increase from $20ms$ and above, however, it does not exceed the expected false-positive proportion until $25ms$ latency is reached.
%TODO: Re-run latency experiment, larger sample size?

@@

As frame-time increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyMid}, however in \ref{fig_RatioVsLatencyHigh} the sample size is small above the tolerance line due to the high occurrences of thrashing and so the relationship in this experiment can not be deduced. For the $8ms$ and $15ms$ experiments the late collisions do not exceed 10\% and are assumed to be accounted for by false-positives. In the $8ms$ frame-time tolerance experiment, late collisions are not expected to occur until $16ms$ frame-time, based on the aura size. Late collisions do not exceed $20\%$ and is assumed to be accounted for by false-positives. In the $15ms$ frame-time tolerance experiment, late collisions exceed $20\%$ once $16ms$ latency is exceeded as predicted by the aura size and increase to over $75\%$ at $30ms$ latency.
In the final frame-time tolerance experiment using $32ms$ frame-time tolerance, late collisions only occur for frame-times of $64ms$. The aura calculation predicts that late collisions should begin to occur as soon as $32ms$ frame-time is exceeded, however, due to the low sample size, as a result of the high occurrences of thrashing, the plot does not accurately represent any relationship between frame-time and proportion of late collisions.

@@

In summary, the varying factors experiments demonstrate that the aura calculation is conclusively correct for speed and frame-time aura tolerance factors and at minimum approximate for latency. For each varied factor, for all three tolerance values chosen, below the tolerance value, the mean + two standard deviations lies under the maximum penetration time expected on a centralised simulation, with only a small number of exceptions discussed above. These experiments also demonstrate the relationship between increasing the tolerance values and amount of thrashing expected to occur. There are no missed collisions, with the exception of the speed experiments and these have significantly fewer than $1\%$ misses below the tolerance and $1.12\%$ misses above the tolerance. Once factors increase above their respective tolerance values, late collisions begin to occur and the more the factors are exceeded, the number and magnitude of errors increase. This demonstrates the correctness of the aura calculation in AP.
%Therefore, we have shown that if the aura calculation factors remain within their respective tolerances, erroneous collisions will be small in magnitude and rare in number. %TODO: Summary Table: Row for each factor: Number of collisions below tolerance, Percentage of errors below tolerance, Mean error below tolerance

@@

For the varying packet-loss experiments, the following parameters were used and output given.
The two servers use the following tolerances: a maximum speed tolerance of $32m\mathord{\cdot}s^{-1}$; maximum latency tolerance of $2ms$; and a maximum frame-time tolerance of $15ms$. Each server has a random delay before starting between $0$ and the target frame-time in order to prevent the two servers always having synchronous update cycles. Each experiment is repeated $50$ times. The mean and $+/-2$ standard deviations of penetration time were plotted against each tolerance factor for each of the three tolerance values chosen. In addition, the ratio of errors to correct collisions and misses to total collisions were plotted against packet-loss.

@@

As packet-loss increases, late and erroneous collisions increase and late collisions increase in magnitude. This is due to messages being delayed as packets are lost and need to be re-sent by RakNet's reliability layer. Missed collisions begin to occur as soon as packet-loss exceeds $1\%$, but increase at a lower rate relative to late collisions. (Note, late collisions begin to occur from $0\%$, this is expected as all factors are set to be equal to their tolerances). These results demonstrate that AP is able to prevent $50\%$ of late collisions even in cases of packet-loss as high as (15\%) and prevent all missed collisions with up to $1\%$ packet-loss, which is much greater packet-loss than would be expected in practice (0.01\%) \cite{ThousandEyesCloudPerf2018}.
